### 2017-04-25 monograph/proposal/methods.tex

\section{Método de pesquisa}

\hspace*{5mm}

        O trabalho...