### 2017-04-25 monograph/proposal/main.tex
Created also the monograph/latex_utilities/commands.tex and
monograph/latex_utilities/basic.tex based on the
monograph/proposal/main.tex preamble.

%
% Licensing
%
%   Copyright 2017 @ Evandro Coan
%
%  This program is free software; you can redistribute it and/or modify it
%  under the terms of the GNU General Public License as published by the
%  Free Software Foundation; either version 3 of the License, or ( at
%  your option ) any later version.
%
%  This program is distributed in the hope that it will be useful, but
%  WITHOUT ANY WARRANTY; without even the implied warranty of
%  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. See the GNU
%  General Public License for more details.
%
%  You should have received a copy of the GNU General Public License
%  along with this program.  If not, see <http://www.gnu.org/licenses/>.
%

%
% Simple Sectioned Essay Template - LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% `proposal.tex`
% Based on http://www.latextemplates.com/template/simple-sectioned-essay
%

%----------------------------------------------------------------------------------------
%   PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

% Sets this document font size to 12pt and set is type to article
\documentclass[12pt]{article}

% https://tex.stackexchange.com/questions/664/why-should-i-use-usepackaget1fontenc
\usepackage[T1]{fontenc}

% Sets this document language usage to Portuguese
\usepackage[brazil]{babel}

% Required to change the page size to A4
% \usepackage{geometry}
%
% Sets this document output papel to A4 within 2 centimeters margin
% \geometry{a4paper}
%
% Set the page size to be A4 as opposed to the default US Letter
\usepackage[a4paper, margin=2cm]{geometry}

% Sets the character set of this document to UTF-8
\usepackage[utf8]{inputenc}

%----------------------------------------------------------------------------------------
%   File settings
%----------------------------------------------------------------------------------------

% Utilities files
\input{../latex_utilities/basic}
\input{../latex_utilities/commands}
\input{../latex_utilities/programming}

% To use the font Times New Roman, instead of the default LaTeX font
% \usepackage{mathptmx}
%
% more up-to-date than 'mathptmx'
\usepackage{newtxtext,newtxmath}

% https://tex.stackexchange.com/questions/10377/texttt-overfull-hbox-problem
% https://tex.stackexchange.com/questions/66052/should-i-load-microtype-with-pdflatex
\usepackage{microtype}

% Indent the first section paragraphs
% https://tex.stackexchange.com/q/39227/119062
\usepackage{indentfirst}

% Line spacing
\linespread{1.5}

% Specifies the directory where pictures are stored
\graphicspath{{pictures/}}

% Underfull \hbox in bibliography
% https://tex.stackexchange.com/questions/10924/underfull-hbox-in-bibliography
\apptocmd{\thebibliography}{\raggedright}{}{}

%----------------------------------------------------------------------------------------
%   DOCUMENT CONTENTS
%----------------------------------------------------------------------------------------

\begin{document}

% pdfTeX warning (ext4): destination with the same identifier (nam e{page.1}) has been already used, duplicate ignored
% https://tex.stackexchange.com/questions/18924/pdftex-warning-ext4-destination-with-the-same-identifier-nam-epage-1-has
\hypersetup{pageanchor=false}

\include{cover}
\include{abstract}
\include{summary}
\include{introduction}
\include{objectives}
\include{methods}
\include{schedule}

\bibliographystyle{abbrv}
\bibliography{refs}

\end{document}