### 2017-04-27 monograph/proposal/main.tex
Replaced the author name hard coded by the macro \Author

%
% Licensing
%
%   Copyright 2017 @ Evandro Coan
%
%  This program is free software; you can redistribute it and/or modify it
%  under the terms of the GNU General Public License as published by the
%  Free Software Foundation; either version 3 of the License, or ( at
%  your option ) any later version.
%
%  This program is distributed in the hope that it will be useful, but
%  WITHOUT ANY WARRANTY; without even the implied warranty of
%  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. See the GNU
%  General Public License for more details.
%
%  You should have received a copy of the GNU General Public License
%  along with this program.  If not, see <http://www.gnu.org/licenses/>.
%

%
% Simple Sectioned Essay Template - LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% `proposal.tex`
% Based on http://www.latextemplates.com/template/simple-sectioned-essay
%

%----------------------------------------------------------------------------------------
%   PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

% Sets this document font size to 12pt and set is type to article
\documentclass[12pt]{article}

% https://tex.stackexchange.com/questions/664/why-should-i-use-usepackaget1fontenc
\usepackage[T1]{fontenc}

% Sets this document language usage to Portuguese
\usepackage[brazil]{babel}

% Required to change the page size to A4
% \usepackage{geometry}
%
% Sets this document output papel to A4 within 2 centimeters margin
% \geometry{a4paper}
%
% Set the page size to be A4 as opposed to the default US Letter
\usepackage[a4paper, margin=2cm]{geometry}

% Sets the character set of this document to UTF-8
\usepackage[utf8]{inputenc}

%----------------------------------------------------------------------------------------
%   File settings
%----------------------------------------------------------------------------------------

% Author name
%
% What do \makeatletter and \makeatother do?
% https://tex.stackexchange.com/questions/8351/what-do-makeatletter-and-makeatother-do
%
% How can I use @author, @date, and @title after maketitle?
% https://tex.stackexchange.com/questions/27710/how-can-i-use-author-date-and-title-after-maketitle
\makeatletter
\author{Evandro Coan}        \let\Author\@author
\makeatother

% Utilities files
\input{../latex_utilities/basic}
\input{../latex_utilities/commands}
\input{../latex_utilities/programming}

% To use the font Times New Roman, instead of the default LaTeX font
% \usepackage{mathptmx}
%
% more up-to-date than 'mathptmx'
\usepackage{newtxtext,newtxmath}

% https://tex.stackexchange.com/questions/10377/texttt-overfull-hbox-problem
% https://tex.stackexchange.com/questions/66052/should-i-load-microtype-with-pdflatex
\usepackage{microtype}

% Indent the first section paragraphs
% https://tex.stackexchange.com/q/39227/119062
\usepackage{indentfirst}

% Why is the linespread factor as it is?
% \linespread{1.5}
% https://tex.stackexchange.com/questions/30073/why-is-the-linespread-factor-as-it-is
\usepackage{setspace}
% \doublespacing
\onehalfspacing

% Specifies the directory where pictures are stored
\graphicspath{{pictures/}}

% Uncomment to remove all indentation from paragraphs
%\setlength\parindent{0pt}

% Bad formatting using URLs in bibtex
% https://tex.stackexchange.com/questions/22888/bad-formatting-using-urls-in-bibtex
\usepackage{etoolbox}

% Underfull \hbox in bibliography
% https://tex.stackexchange.com/questions/10924/underfull-hbox-in-bibliography
\apptocmd{\thebibliography}{\raggedright}{}{}

% How to avoid overfull error with url package?
% See also the `\usepackage{url}` declarationon the file `basic.tex`.
% Set this to 2mu or 3mu if URL start troubling again.
% https://tex.stackexchange.com/questions/261776/how-to-avoid-overfull-error-with-url-package
\Urlmuskip=0mu plus 1mu

% Automatically put a `\medskip` spacing between paragraphs
% https://tex.stackexchange.com/questions/365976/how-to-stop-the-package-usepackageparskip-disabling-the-paragraph-indentation
% \edef\restoreparindent{\parindent=\the\parindent\relax}
% \usepackage{parskip}
% \restoreparindent
%
% How to restore the parskip skips between list items?
% https://tex.stackexchange.com/questions/366848/how-to-restore-the-parskip-skips-between-list-items
\parskip=0.5\baselineskip \advance\parskip by 0pt plus 2pt

%----------------------------------------------------------------------------------------
%   DOCUMENT CONTENTS
%----------------------------------------------------------------------------------------

\begin{document}

% pdfTeX warning (ext4): destination with the same identifier (nam e{page.1}) has been already used, duplicate ignored
% https://tex.stackexchange.com/questions/18924/pdftex-warning-ext4-destination-with-the-same-identifier-nam-epage-1-has
\hypersetup{pageanchor=false}

\include{cover}
\include{abstract}
\include{summary}
\include{introduction}
\include{objectives}
\include{methods}
\include{schedule}

\bibliographystyle{abbrv}
\bibliography{refs}

\end{document}