### 2017-08-26 monograph/thesis/chapters/chapter_1.tex
Added the ufscthesis template files to the main document main.tex

\cleardoublepage
\phantomsection %The \phantomsection command is needed to create a link to a place in the document that is not a figure, equation, table, section, subsection, chapter, etc.
\addcontentsline{toc}{chapter}{\texorpdfstring{\MakeTextUppercase{Capítulo 3}}{Capítulo 3}}

\chapter{Lorem ipsum dolor sit amet}

\begin{flushright}
    \showfont
\end{flushright}

\newpage

% ---
\section{Aliquam vestibulum fringilla lorem}
% ---

Lipsum me [1]

Lipsum me [2-3]