### 2017-08-28 monograph/thesis/pretexto/fichacatalografica.tex
Created the fichacatalografica.tex no warranty notice.

\ifenglish

Legal Notes

There is no warranty for any part of the documented software. The authors have taken care in the
preparation of this thesis, but make no expressed or implied warranty of any kind and assume no
responsibility for errors or omissions. No liability is assumed for incidental or consequential
damages in connection with or arising out of the use of the information or programs contained here.
\cite{koma-scrguien}

\else

Notas legais

Não há garantia para qualquer parte do software documentado. Os autores tomaram cuidado na
preparação desta tese, mas não fazem nenhuma garantia expressa ou implícita de qualquer tipo e não
assumem qualquer responsabilidade por erros ou omissões. Não se assume qualquer responsabilidade por
danos incidentais ou consequentes em conexão ou decorrentes do uso das informações ou programas aqui
contidos. \cite{koma-scrguien}

\fi

% http://portalbu.ufsc.br/ficha
% http://portal.bu.ufsc.br/servicos/ficha-de-identificacao-da-obra/
\begin{fichacatalografica}
    \vspace*{\fill}

    \begin{center}

        Catalogação na fonte pela Biblioteca Universitária da Universidade Federal de Santa Catarina.

        Arquivo compilado às \currenttime h do dia \today.

        \framebox[\textwidth]
        {
            \begin{minipage}{0.98\textwidth}

                \ttfamily
                \imprimirautor

                \hspace{0.5cm} \imprimirtitulo~:~\imprimirsubtitulo~/~\imprimirautor;
                orientador(a),~\imprimirorientador;~co\hyp{}orientador(a),~\imprimircoorientador
                ~--~\imprimirlocal,~\currenttime,~\imprimirdata.

                % Prints how much pages there are on the document and links to the last page
                \hspace{0.5cm} \pageref{LastPage} p.
                \bigskip

                \hspace{0.5cm} \imprimirtipotrabalho~--~\imprimirinstituicao,
                \imprimircentro,~\imprimirprograma.
                \bigskip

                \hspace{0.5cm} Inclui referências
                \bigskip

                \hspace{0.5cm}
                    1. Uma Palavra\hyp{}chave ~
                    2. Outra Palavra\hyp{}chave ~
                    3. Mais Palavras\hyp{}chave ~
                    I. \imprimirorientador ~
                    II. \imprimircoorientador ~
                    III. \imprimirprograma ~
                    IV. \imprimirtitulo ~
                \bigskip

                \hspace{7.75cm} CDU 02:141:005.7

            \end{minipage}
        }

    \end{center}

\end{fichacatalografica}