### 2017-08-30 monograph/thesis/pretexto/folhadeaprovacao.tex

\addtotextpreliminarycontent{\chooselang{Approval Sheet}{Folha de Aprovação}}

\begin{folhadeaprovacao}

    \begin{center}
        {\ABNTEXchapterfont\large\imprimirautor}

        \begin{center}
            \ABNTEXchapterfont\bfseries\Large\imprimirtitulo
        \end{center}

        \begin{minipage}{\textwidth}
            \chooselang
            {
                This thesis was considered appropriate to obtain the Doctor 's Degree in Electrical Engineering, in the area of concentration in Power Electronics and Electrical Drive, and approved in its final form by the Post-Graduate Program in Electrical Engineering of the Federal University of Santa Catarina.
            }
            {
                Esta Tese foi julgada adequada para obtenção do Título de Doutor em Engenharia Elétrica, na área de concentração em Eletrônica de Potência e Acionamento Elétrico, e aprovada em sua forma final pelo Programa de Pós--Graduação em Engenharia Elétrica da Universidade Federal de Santa Catarina.
            }
        \end{minipage}%

    \end{center}
    \begin{center}
        Florianópolis, \imprimirdata.
    \end{center}

    \assinatura
    {
        \textbf{Prof. Marcelo Lobo Heldwein, Dr.} \\
        \chooselang{Coordinator of the}{Coordenador do} \imprimirprograma
    }

    \assinatura
    {
        \textbf{\imprimirorientador} \\ \imprimirorientadorRotulo \\
        \imprimirinstituicao~--~\imprimirinstituicaosigla
    }

    \assinatura
    {
        \textbf{\imprimircoorientador} \\ \imprimircoorientadorRotulo \\
        \imprimirinstituicao~--~\imprimirinstituicaosigla
    }

    \newpage
    \begin{flushleft}
        \textbf{\chooselang{Examination Board:}{Banca Examinadora:}}
    \end{flushleft}

    \assinatura{\textbf{Prof. Arnaldo José Perin, \chooselang{PhD.}{Dr.}} \\
    \imprimirinstituicao~--~\imprimirinstituicaosigla}

    \assinatura{\textbf{Prof. Denizar Cruz Martins, \chooselang{PhD.}{Dr.}} \\
    \imprimirinstituicao~--~\imprimirinstituicaosigla}

    \assinatura{\textbf{Prof. Roberto Francisco Coelho, \chooselang{PhD.}{Dr.}} \\
    \imprimirinstituicao~--~\imprimirinstituicaosigla}

    \assinatura{\textbf{Prof. Samir Ahmad Mussa, \chooselang{PhD.}{Dr.}} \\
    \imprimirinstituicao~--~\imprimirinstituicaosigla}

    \assinatura{\textbf{Prof. Telles Brunelli Lazzarin, \chooselang{PhD.}{Dr.}} \\
    \imprimirinstituicao~--~\imprimirinstituicaosigla}

\end{folhadeaprovacao}