### 2017-09-12 monograph/thesis/pretexto/folhadeaprovacao.tex

\addtotextpreliminarycontent{\lang{Approval Sheet}{Folha de Aprovação}}

\begin{folhadeaprovacao}

    \begin{center}
        {\ABNTEXchapterfont\large\imprimirautor}

        \begin{center}
            \ABNTEXchapterfont\bfseries\Large\imprimirtitulo
        \end{center}

        \begin{minipage}{\textwidth}
            \lang
            {
                This thesis was considered appropriate to obtain the \imprimirformacao, in the area
                of \imprimirarea, and approved in its final form by the \imprimirprograma of the
                \imprimirinstituicao.
            }
            {
                Esta Tese foi julgada adequada para obtenção do Título de \imprimirformacao, na área
                de \imprimirarea, e aprovada em sua forma final pelo \imprimirprograma da
                \imprimirinstituicao.
            }
        \end{minipage}%

    \end{center}
    \begin{center}
        Florianópolis, \imprimirdata.
    \end{center}

    \assinatura{\textbf{Prof. Renato Cislaghi, \lang{PhD.}{Dr.}} \\
    \lang{Coordinator of the}{Coordenador do} \imprimirprograma}

    \assinatura{\textbf{\imprimirorientador} \\ \imprimirorientadorRotulo \\
    \imprimirinstituicao~--~\imprimirinstituicaosigla}

    \assinatura{\textbf{\imprimircoorientador} \\ \imprimircoorientadorRotulo \\
    \imprimirinstituicao~--~\imprimirinstituicaosigla}

    \newpage
    \begin{flushleft}
        \textbf{\lang{Examination Board:}{Banca Examinadora:}}
    \end{flushleft}

    \assinatura{\textbf{Prof. Jerusa Marchi, \lang{PhD.}{Dr.}} \\
    \imprimirinstituicao~--~\imprimirinstituicaosigla}

    \assinatura{\textbf{Prof. Elder Rizzon Santos, \lang{PhD.}{Dr.}} \\
    \imprimirinstituicao~--~\imprimirinstituicaosigla}

\end{folhadeaprovacao}