### 2017-09-15 monograph/thesis/postexto/abstracts.tex
Created the annex postexto/abstracts.tex to save the references'
abstracts.

% The \phantomsection command is needed to create a link to a place in the document that is not a
% figure, equation, table, section, subsection, chapter, etc.
%
% When do I need to invoke \phantomsection?
% https://tex.stackexchange.com/questions/44088/when-do-i-need-to-invoke-phantomsection
\cleardoublepage
\phantomsection

% Is it possible to keep my translation together with original text?
% https://tex.stackexchange.com/questions/5076/is-it-possible-to-keep-my-translation-together-with-original-text
\chapter{\lang{References' Abstracts}{Resumos das Referências}}

\begin{englishtext}
\begin{enumerate}

    \item In this paper we argue that there is a necessity for automating modifications to legacy assets.
    We propose a five layered process for the introduction and employment of tool support that
    enables automated modification to entire legacy systems. Furthermore, we elaborately discuss
    each layer on a conceptual level, and we make appropriate references to sources where technical
    contributions supporting that particular layer can be found. We sketch the perspective that more
    and more people working in the software engineering area will be contributing to working on
    existing systems and/or tools to support such work. \cite{legacyAssets}

\end{enumerate}
\end{englishtext}