### 2017-09-16 monograph/thesis/postexto/abstracts.tex
Added all references.bib abstracts on abstracts.tex

% The \phantomsection command is needed to create a link to a place in the document that is not a
% figure, equation, table, section, subsection, chapter, etc.
%
% When do I need to invoke \phantomsection?
% https://tex.stackexchange.com/questions/44088/when-do-i-need-to-invoke-phantomsection
\cleardoublepage
\phantomsection

% Is it possible to keep my translation together with original text?
% https://tex.stackexchange.com/questions/5076/is-it-possible-to-keep-my-translation-together-with-original-text
\chapter{\lang{References' Abstracts}{Resumos das Referências}}

\begin{englishtext}
\begin{enumerate}

    \item \textbf{The Guide KOMA - Script}: KOMA-Script  is very complex. This is
    evidenced by the fact that it consists of not only a single class or a
    single package, but a bundle of many classes and packages. Although the
    classes are designed as a counterpart to the standard classes, that does not
    necessarily mean that they only have the commands, environments, and setting
    of the standard classes or imitate their appearance. The capabilities of
    KOMA-Script surpass the capabilities of the standard classes considerably.
    Some of them are to be regarded as a supplement to the basic skills of the
    LATEX kernel. \cite{komaScrguien}

    \item \textbf{Death to the Space Infidels!}: Ah, spring. What a wonderful time of
    year. A time when young programmers' minds turn to thoughts of ... never
    ending last-man-standing filibuster arguments about code formatting.
    Naturally. And there is no argument more evergreen than the timeless debate
    between tabs and spaces. \cite{Atwood}

    \item \textbf{Is imposing the same code format for all developers a good idea?}: We
    are considering to impose a single standard code format in our project (auto
    format with save actions in Eclipse). The reason is that currently there is
    a big difference in the code formats used by several (>10) developers which
    makes it harder for one developer to work on the code of another developer.
    The same Java file sometimes uses 3 different formats. \cite{Geukens}

    \item \textbf{Powerful code indenter front-end, UniversalIndentGUI}: Ever concerned
    about how your code looks like? Ever heard of different indenting styles,
    for example K\&R? Ever received code from someone else who didn't care about
    code formatting? Ever tried to configure a code indenter to convert such
    code to your coding style? Ever got bored by that tedious "changing a
    parameter"-"call the indeter"-"try and error" procedure? Help is close to
    you. UniversalIndentGUI offers a live preview for setting the parameters of
    nearly any indenter. You change the value of a parameter and directly see
    how your reformatted code will look like. Save your beauty looking code or
    create an anywhere usable batch/shell script to reformat whole directories
    or just one file even out of the editor of your choice that supports
    external tool calls.

    \item \textbf{SUBLIME TEXT 3 DOCUMENTATION, Syntax Definitions}: Sublime Text can
    use both .sublime-syntax and .tmLanguage files for syntax highlighting. This
    document describes .sublime-syntax files. Sublime Syntax files are YAML
    files with a small header, followed by a list of contexts. Each context has
    a list of patterns that describe how to highlight text in that context, and
    how to change the current text. \cite{Skinner}

    \item \textbf{Towards a Universal Code Formatter through Machine Learning}: There
    are many declarative frameworks that allow us to implement code formatters
    relatively easily for any specific language, but constructing them is
    cumbersome. The first problem is that “everybody” wants to format their code
    differently, leading to either many formatter variants or a ridiculous
    number of configuration options. Second, the size of each implementation
    scales with a language’s grammar size, leading to hundreds of rules. In this
    paper, we solve the formatter construction problem using a novel approach,
    one that automatically derives formatters for any given language without
    intervention from a language expert. We introduce a code formatter called
    CODEB UFF that uses machine learning to abstract formatting rules from a
    representative corpus, using a carefully designed feature set. Our
    experiments on Java, SQL, and ANTLR grammars show that CODEB UFF is
    efficient, has excellent accuracy, and is grammar invariant for a given
    language. It also generalizes to a 4th language tested during manuscript
    preparation. \cite{universalCodeFormatter}

    \item \textbf{An industrial application of context-sensitive formatting}: Automated
    formatting is an important technique for the software maintainer. It is
    either applied separately to improve the readability of source code, or as
    part of a source code transformation tool chain. In this paper we report on
    the application of generic tools for constructing formatters. In an
    industrial setting automated formatters need to be tailored to the
    requirements of the customer. The (legacy) programming language or dialect
    and the corporate formatting conventions are specific and non-negotiable.
    Can generic formatting tools deal with such unexpected requirements? Driven
    by an industrial case of 78 thousand lines of Cobol code, several
    limitations in existing formatting technology have been addressed. We
    improved its flexibility by replacing a generative phase by a generic tool,
    and we added a little expressiveness to the formatting backend. Most
    importantly, we employed a multi-stage formatting architecture that can cope
    with any kind of formatting convention using more computational power.
    \cite{industrialApplication}

    \item \textbf{Program indentation and comprehensibility}: The consensus in the
    programming community is that indentation aids program comprehension,
    although many studies do not back this up. We tested program comprehension
    on a Pascal program. Two styles of indentation were used -- blocked and
    non-blocked -- in addition to four passible levels of indentation
    (0, 2, 4, 6 spaces). Both experienced and novice subjects were used.
    Although the blocking style made no difference, the level of indentation had
    a significant effect on program comprehension. (2--4 spaces had the highest
    mean score for program comprehension.) We recommend that a moderate level
    of indentation be used to increase program comprehension and user
    satisfaction. \cite{programIndentation}

    \item \textbf{A Language Independent Framework for Context-sensitive Formatting}:
    Automated formatting is an important technique for the software maintainer.
    It is either applied separately to improve the readability of source code,
    or as part of a source code transformation tool chain. In this paper we
    report on the application of generic tools for constructing formatters. In
    an industrial setting automated formatters need to be tailored to the
    requirements of the customer. The (legacy) programming language or dialect
    and the corporate formatting conventions are specific and non-negotiable.
    Can generic formatting tools deal with such unexpected requirements? Driven
    by an industrial case of nearly 80 thousand lines of Cobol code, several
    limitations in existing formatting technology have been addressed. We
    improved its flexibility by replacing a generative phase by a generic tool,
    and we added a little expressiveness to the formatting back end. Most
    importantly, we employed a multi-stage formatting framework that can cope
    with any kind of formatting convention using more computational power.
    \cite{independentFramework}

    \item \textbf{An architecture for context-sensitive formatting}: We developed an
    architecture for context-sensitive formatting of source code. The
    architecture was implemented and applied in an industrial formatting case.
    \cite{architectureFormatting}

    \item \textbf{Pretty-printing for software reengineering}: Automatic software
    reengineerings change or repair existing software systems. They are usually
    tailor-made for a specific customer and language dependent. Maintaining
    similar reengineerings for multiple customers and different language
    dialects might therefore soon become problematic unless advanced language
    technology is being used. Generic pretty-printing is part of such technology
    and is the subject of this paper. We discuss specific pretty-print aspects
    of software reengineering such as fulfilling customer-specific format
    conventions, preserving existing layout, and producing multiple output
    formats. In addition, we describe pretty-print techniques that help to
    reduce maintenance effort of tailor-made reengineerings supporting multiple
    language dialects. Applications, such as COBOL reengineering and SDL
    documentation generation show that our techniques, implemented in the
    generic pretty-printer GPP, are feasible. \cite{prettyPrinting}

    \item \textbf{Automated Mass Maintenance of Software Assets}: This is a research
    summary of a PhD project in the area of massive software maintenance
    automation. We explain the context, approach, and contributions.
    \cite{massMaintenance}

    \item \textbf{Towards automated modification of legacy assets}: In this paper we
    argue that there is a necessity for automating modifications to legacy
    assets. We propose a five layered process for the introduction and
    employment of tool support that enables automated modification to entire
    legacy systems. Furthermore, we elaborately discuss each layer on a
    conceptual level, and we make appropriate references to sources where
    technical contributions supporting that particular layer can be found. We
    sketch the perspective that more and more people working in the software
    engineering area will be contributing to working on existing systems and/or
    tools to support such work. \cite{legacyAssets}

    \item \textbf{Automated maintenance of a software portfolio}: This is an experience
    report on automated mass maintenance of a large Cobol software portfolio. A
    company in the financial services and insurance industry upgraded their
    database system to a new version, affecting their entire software portfolio.
    The database system was accessed by the portfolio of 45 systems, totalling
    nearly 3000 programs and covering over 4 million lines of Cobol code. We
    upgraded the programs to the new database version using several automatic
    tools, and we performed an automated analysis supporting further manual
    modifications by the system experts. The automatic tools were built using a
    combination of lexical and syntactic technology, and they were deployed in a
    mass update factory to allow large-scale application to the software
    portfolio. The updated portfolio has been accepted and taken into production
    by the company, serving over 600 employees with the new database version. In
    this paper, we discuss the automated upgrade from problem statement to
    project costs. \cite{softwarePortfolio}

    \item \textbf{2-Head Pushdown Automata}: Finite state automata recognize regular
    languages which can be used in text processing, compilers, and hardware
    design. Two head finite automata accept linear context free languages. In
    addition, pushdown automata are able to recognize context free languages
    which can be used in programming languages and artificial intelligence. The
    finite automaton has deterministic and non-deterministic version likewise
    the two head finite automata and the pushdown automata. The deterministic
    version of these machines is such that there is no choice of move in any
    situation while the non-deterministic version has a choice of move. In this
    research the 2-head pushdown automata are described which is more powerful
    than the pushdown automata and it is able to recognize some non-context free
    languages as well. During this work, the main task is to characterize these
    machines. \cite{pushdownAutomata}

    \item \textbf{Aspect-oriented model-driven code generation: A systematic mapping
    study}: Context: Model-driven code generation is being increasingly applied
    to enhance software development from perspectives of maintainability,
    extensibility and reusability. However, aspect-oriented code generation from
    models is an area that is currently underdeveloped. Objective: In this study
    we provide a survey of existing research on aspect-oriented modeling and
    code generation to discover current work and identify needs for future
    research. Method: A systematic mapping study was performed to find relevant
    studies. Classification schemes have been defined and the 65 selected
    primary studies have been classified on the basis of research focus,
    contribution type and research type. Results: The papers of solution
    proposal research type are in a majority. All together aspect-oriented
    modeling appears being the most focused area divided into modeling notations
    and process (36\%) and model composition and interaction management (26\%).
    The majority of contributions are methods. Conclusion: Aspect-oriented
    modeling and composition mechanisms have been significantly discussed in
    existing literature while more research is needed in the area of
    model-driven code generation. Furthermore, we have observed that previous
    research has frequently focused on proposing solutions and thus there is
    need for research that validates and evaluates the existing proposals in
    order to provide firm foundations for aspect-oriented model-driven code
    generation. \cite{aspectOriented}

    \item \textbf{A Review: Analysis of Aspect Orientation and Model Driven Engineering
    for Code Generation}: In the development of large and complex software
    application software engineers has to focuses on many requirements other
    than desired application’s requirement at the coding and design level. Code
    Generation is a technique which is use to automatically generates lower
    level executable code from higher level design artifact. Code generation
    provides design of the code at higher abstract level so that software
    developers can focuses on higher level design problem simultaneously meeting
    goals of desired application. Aspect Orientation (AO) is characterizing by
    identification and separation of different concerns and encapsulates them in
    modules. Concern is an interest which pertains to system operation,
    function, development or any other things which is important to one of the
    stakeholder. Model Driven Engineering (MDE) is a development paradigm which
    is characterize by model transformation and uses models to support various
    stages of the development life cycle. Model is primary artifact in MDE. In
    this paper we analyze both techniques i.e. AO and MDE, and how they can be
    used for code generation. \cite{aspectOrientationReview}

\end{enumerate}
\end{englishtext}