### 2017-11-04 monograph/thesis/chapters/conclusion.tex
Also added the go to top/back on the memoir's chapters titles

% The \phantomsection command is needed to create a link to a place in the document that is not a
% figure, equation, table, section, subsection, chapter, etc.
%
% When do I need to invoke \phantomsection?
% https://tex.stackexchange.com/questions/44088/when-do-i-need-to-invoke-phantomsection
\phantomsection

% ---
% Conclusão (outro exemplo de capítulo sem numeração e presente no sumário)
% ---
\addcontentsline{toc}{chapter}{\texorpdfstring{\MakeTextUppercase{Conclusion}}{Conclusion}}

\chapter*[Conclusion]{Conclusion}

    Lipsum me [31-33]