### 2017-11-05 monograph/thesis/chapters/conclusion.tex
Fixed the conclusion chapter showing together with the last chapter

% The \phantomsection command is needed to create a link to a place in the document that is not a
% figure, equation, table, section, subsection, chapter, etc.
%
% When do I need to invoke \phantomsection?
% https://tex.stackexchange.com/questions/44088/when-do-i-need-to-invoke-phantomsection
\phantomsection

% ---
% Conclusão (outro exemplo de capítulo sem numeração e presente no sumário)
% ---
\chapter[]{\lang{Conclusion}{Conclusão}}

    Lipsum me [31-33]