### 2017-11-06 monograph/thesis/chapters/motivation.tex
Created the Code Beautifying section on motivation.tex

% The \phantomsection command is needed to create a link to a place in the document that is not a
% figure, equation, table, section, subsection, chapter, etc.
%
% When do I need to invoke \phantomsection?
% https://tex.stackexchange.com/questions/44088/when-do-i-need-to-invoke-phantomsection
\phantomsection

% Is it possible to keep my translation together with original text?
% https://tex.stackexchange.com/questions/5076/is-it-possible-to-keep-my-translation-together-with-original-text
\chapter{\lang{Motivation}{Motivação}}

    \section{What does coding is?}

    Coding is like writing and reading a book for the large people, you like it
    to look beautifully. Or at least do you expect such when you buy a book, for
    example, to learn programming for you first time \cite{howNovicesRead}.
    You expect:

    \begin{enumerate}
        \item Things to be well organized, so you do not get lost.
        \item The colors to be properly placed, so you do not get distracted
              from the main content.
        \item The spacing between paragraphs, words, chapters, sections
              subsections, etc, to be well adjusted. Not everything cluttered in
              only one file, line, function, class, or whatsoever so.
    \end{enumerate}

    \section{Computer Assisted Programming}

    Your computer should help you with with these unforeseen tasks. Why should I
    spend my precious time checking whether I am actually copying something
    space indented, when I am actually coping something tab indented?

    Therefore, how to do such a thing on this 21\q{}st century? Perhaps we
    should sit and cry while waiting for some greater force to come and rescue
    us. Or may be you should stop crying and actually do something about other
    than keep waiting for you mommy to come and save you from the darkness
    growing behind you back leading you to endless unsleepy nights fixing your
    code just because everything just went wrong.

    \section{How to Write Good Code}

    If you ask around, you will probably listen you need to write code to learn
    how to write good code. This seems to make sense like training for the next
    Olympics. As long as you train very hard, you get results.

    However, the same way as for gymnasium training there is the right or
    perhaps nice way to do it and the wrong way, for computer programming is the
    same.

    \section{Code Beautifying}

    A robust Code Beautifier can get a lot more complicated just with the basic
    definitions of formatting (\nameref{source_code_beautifiers})
    applied over each language own characteristics
    as for example:

    \begin{enumerate}[nosep,nolistsep]
        \item Add spaces before if\s name as in `if(var)' versus `if (var)'
        \item Add spaces inside if\s as in `if(var)' versus `if( var )'
        \item Add spaces before for\s name as in `for(var)' versus `for (var)'
        \item Add spaces inside for\s as in `for(var)' versus `for( var )'
        \item ...
    \end{enumerate}

    As may be noticed, the list may became quite big, and if fact such big list
    of rules has been implemented.
    Looking over the Beautifier called `Uncrustify'\footnote{\url{https://github.com/uncrustify/uncrustify}},
    we can find about 500 settings with specifications like these above.