### 2017-12-03 monograph/proposal/main.tex
Added \raggedbottom to monograph/proposal/main.tex and updated the
monograph/proposal/utilities submodule.

%
% Simple Sectioned Essay Template - LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% `proposal.tex`
% Based on:
%
% 1. https://github.com/royertiago/tcc
% 2. http://www.latextemplates.com/template/simple-sectioned-essay
%

%----------------------------------------------------------------------------------------
%   PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

% Sets this document font size to 12pt and set is type to article
\documentclass[12pt]{article}

% Sets this document language usage to Portuguese
\usepackage[brazil]{babel}

% Sets this document output papel to A4 within 2 centimeters margin
% \PassOptionsToPackage{showframe}{geometry}
\usepackage[a4paper,margin=2cm]{geometry}

%----------------------------------------------------------------------------------------
%   File settings
%----------------------------------------------------------------------------------------

% Utilities files
\input{utilities/utilities}

% To use the font Times New Roman, instead of the default LaTeX font
% \usepackage{mathptmx}
%
% more up-to-date than 'mathptmx'
\usepackage{newtxtext,newtxmath}

% Always use it as should improve full justification
% https://tex.stackexchange.com/questions/10377/texttt-overfull-hbox-problem
% https://tex.stackexchange.com/questions/66052/should-i-load-microtype-with-pdflatex
\usepackage{microtype}

% Indent the first section paragraphs
% https://tex.stackexchange.com/q/39227/119062
\usepackage{indentfirst}

% Why is the linespread factor as it is?
% \linespread{1.5}
% https://tex.stackexchange.com/questions/30073/why-is-the-linespread-factor-as-it-is
\usepackage{setspace}
% \doublespacing
\onehalfspacing

% Specifies the directory where pictures are stored
\graphicspath{{pictures/}}

%----------------------------------------------------------------------------------------
%   DOCUMENT CONTENTS
%----------------------------------------------------------------------------------------

% Author name
%
% What do \makeatletter and \makeatother do?
% https://tex.stackexchange.com/questions/8351/what-do-makeatletter-and-makeatother-do
%
% How can I use @author, @date, and @title after maketitle?
% https://tex.stackexchange.com/questions/27710/how-can-i-use-author-date-and-title-after-maketitle
\makeatletter
\author{Evandro Coan}\let\Author\@author
% \date{Somewhen}           \let\Date\@date
\def\Advisor{Thiago Ângelo Gelaim}
\def\Supervisor{Ricardo Azambuja Silveira}
\makeatother

\begin{document}

% underfull vbox (badness 10000) has occurred while \output is active with memoir
% https://tex.stackexchange.com/questions/65711/underfull-vbox-badness-10000-with-memoir
\raggedbottom

% pdfTeX warning (ext4): destination with the same identifier (nam e{page.1}) has been already used, duplicate ignored
% https://tex.stackexchange.com/questions/18924/pdftex-warning-ext4-destination-with-the-same-identifier-nam-epage-1-has
\hypersetup{pageanchor=false}

\include{cover}
\include{abstract}
\include{summary}

\hypersetup{pageanchor=true}

% To automatically put a [Go To Top] on each section
\addGoToSummary

\include{introduction}
\include{schedule}
\include{implementation}

\bibliographystyle{abbrv}
\bibliography{refs}

\end{document}